---
title: "Untitled"
author: "Erin M. Buchanan"
date: "3/5/2018"
output: pdf_document
---

# Experiment 2

# Method

## Participants

```{r datascreen2, include = FALSE}

exp2data = read.csv("exp2_final.csv")

####data screening####
##accuracy checks
summary(exp2data)
#table(exp2data$partno)
#table(exp2data$speed)
#hist(exp2data$speed)
exp2data$realfake = factor(exp2data$realfake,
                            levels = c("fake", "real"))
part2 = length(table(exp2data$partno))
#get rid of blank row at the end
exp2data = na.omit(exp2data)

##missing data
##no missing data to replace
##figure out the couple missing data points
rows2 = nrow(exp2data)

##outliers
#with(exp2data, plot(speed, rating))
mahal = mahalanobis(exp2data[ , c(5,7)],
                    colMeans(exp2data[ , c(5,7)], na.rm = T),
                    cov(exp2data[ , c(5,7)], use = "pairwise.complete.obs"))
cutoff = qchisq(1-.001, ncol(exp2data[ , c(5,7)]))
table(mahal < cutoff)
#subset(exp2data, mahal > cutoff)

##assumptions
correlations = cor(exp1data[, c(5,7, 13, 14, 15, 17)], use="pairwise.complete.obs")
correlations
symnum(correlations)

##make the random stuff
random = rchisq(nrow(exp2data), 7)
#fake = lm(random~., data=exp1data[ , c(5, 7, 12, 13, 14, 15, 17)])
fake = lm(random~., data=exp2data[ , c(5, 7)])

##get the linearity plot
##create the standardized residuals
standardized = rstudent(fake)
{qqnorm(standardized)
abline(0,1)}

##multivariate normality
hist(standardized, breaks=15)

##homogeneity and homoscedaticity
fitvalues = scale(fake$fitted.values)
{plot(fitvalues, standardized) 
abline(0,0)
abline(v = 0)}

exp2data$hand[exp2data$hand == "left"] = "Left"
exp2data$hand[exp2data$hand == "right"] = "Right"
exp2data$hand = droplevels(exp2data$hand)

temp2 = exp2data[!duplicated(exp2data$partno), ]
hand2 = table(temp2$hand) / length(temp2$hand) * 100

```

Similar to Experiment 1, `r part2` participants were recruited from the university undergraduate human subject pool and received course credit for their time. `r rows2` rows of data were present for these participants, and no data was missing. Rating data were screened for multivariate outliers. Again, part of one participant's ratings were found to have extreme Mahalanobis distance scores [@Tabachnick2012]. However, this individual's ratings were left in the data set. Approximately `r apa(hand2['Left'], 1)` percent of the sample was left-handed. The average typing speed was `r apa(mean(temp2$speed), 2)` (*SD* = `r apa(sd(temp2$speed), 2)`), and the average percent accuracy rate for the typing test was `r apa(mean(temp2$error), 2)` (*SD* = `r apa(sd(temp2$error), 2)`).

## Materials

In this experiment, a smaller subset of words (`r length(table(droplevels(exp2data$trialcode)))`) from Experiment 1 were used, which were split evenly between pseudowords and real words. Average word length was `r apa(mean(exp2data$length), 2)` (*SD* = `r apa(sd(exp2data$length), 2)`; range = `r min(exp2data$length)` - `r max(exp2data$length)`).  

## Procedure

In this study, when participants were shown the word (or pseudoword) on the screen, they were first asked to type the word on the keyboard in front of them. After they had typed the word, they were then asked to rate the word for pleasantness using the scale and self-assessment manikin discussed previously.  

# Results

```{r exp2setup, include = FALSE}
####separate real fake####
real = subset(exp2data,realfake == "real")
fake = subset(exp2data,realfake == "fake")

####centering the variables####
real$speed = scale(real$speed, scale = F)
real$lrswitch = scale(real$lrswitch, scale = F)
real$fingerswitch = scale(real$fingerswitch, scale = F)
real$rha = scale(real$rha, scale = F)

fake$speed = scale(fake$speed, scale = F)
fake$lrswitch = scale(fake$lrswitch, scale = F)
fake$fingerswitch = scale(fake$fingerswitch, scale = F)
fake$rha = scale(fake$rha, scale = F)

```

## Main Effects

```{r meexp2, include = FALSE}

####intercept only model####
model1f = gls(rating ~ 1, 
             data = fake, 
             method = "ML", 
             na.action = "na.omit")
#summary(model1f)

model1r = gls(rating ~ 1, 
             data = real, 
             method = "ML", 
             na.action = "na.omit")
#summary(model1r)

####random intercept only model####
model2f = lme(rating ~ 1, 
             data = fake, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
#summary(model2f)

model2r = lme(rating ~ 1, 
             data = real, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
#summary(model2r)

####controls predictor models####
model3f = lme(rating ~ letterfreq, 
              data = fake, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
#summary(model3f)

model3r = lme(rating ~ letterfreq, 
              data = real, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
#summary(model3r)

####main effects models####
model4f = lme(rating ~ letterfreq + speed + lrswitch + fingerswitch + rha, 
              data = fake, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
#summary(model4f)

model4r = lme(rating ~ letterfreq + speed + lrswitch + fingerswitch + rha, 
              data = real, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
#summary(model4r)

```

```{r model-table2, echo = FALSE, results='asis'}

####interaction models####
model5f = lme(rating ~ letterfreq + speed*lrswitch*fingerswitch*rha, 
              data = fake, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
#summary(model5f)

model5r = lme(rating ~ letterfreq + speed*lrswitch*fingerswitch*rha, 
              data = real,
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
#summary(model5r)

##make the excel sheet
excelprint = matrix(NA, nrow = 54, ncol = 8)
colnames(excelprint) = c("Word Type", "Model", "Predictor", "Estimate", "SE","df", "t", "p")

excelprint[1, 1:3] = c("Psuedo", "Intercept Only",  rownames(summary(model1f)$tTable))
excelprint[1, 4:8] = c(summary(model1f)$tTable[1:2], 32541, summary(model1f)$tTable[3:4])
excelprint[2, 1:3] = c("Psuedo", "Random Intercept",  rownames(summary(model2f)$tTable))
excelprint[2, 4:8] = summary(model2f)$tTable
excelprint[3, 1:3] = c("Psuedo", "Adjustor Variable",  rownames(summary(model3f)$tTable)[1])
excelprint[4, 1:3] = c("Psuedo", "Adjustor Variable",  rownames(summary(model3f)$tTable)[2])
excelprint[3:4, 4:8] = summary(model3f)$tTable
excelprint[5:10 , 1] = c("Psuedo")
excelprint[5:10 , 2] = c("Main Effects")
excelprint[5:10, 3] = rownames(summary(model4f)$tTable)
excelprint[5:10, 4:8] = summary(model4f)$tTable
excelprint[11:27, 1] = c("Psuedo")
excelprint[11:27, 2] = c("Interactions")
excelprint[11:27, 3] = rownames(summary(model5f)$tTable)
excelprint[11:27, 4:8] = summary(model5f)$tTable

excelprint[28, 1:3] = c("Real", "Intercept Only",  rownames(summary(model1r)$tTable))
excelprint[28, 4:8] = c(summary(model1r)$tTable[1:2], 32690, summary(model1r)$tTable[3:4])
excelprint[29, 1:3] = c("Real", "Random Intercept",  rownames(summary(model2r)$tTable))
excelprint[29, 4:8] = summary(model2r)$tTable
excelprint[30, 1:3] = c("Real", "Adjustor Variable",  rownames(summary(model3r)$tTable)[1])
excelprint[31, 1:3] = c("Real", "Adjustor Variable",  rownames(summary(model3r)$tTable)[2])
excelprint[30:31, 4:8] = summary(model3r)$tTable
excelprint[32:37 , 1] = c("Real")
excelprint[32:37 , 2] = c("Main Effects")
excelprint[32:37, 3] = rownames(summary(model4r)$tTable)
excelprint[32:37, 4:8] = summary(model4r)$tTable
excelprint[38:54, 1] = c("Real")
excelprint[38:54, 2] = c("Interactions")
excelprint[38:54, 3] = rownames(summary(model5r)$tTable)
excelprint[38:54, 4:8] = summary(model5r)$tTable

write.csv(excelprint, "Predictor Statistics Exp 2.csv", row.names = F)

faketable = anova(model1f, model2f, model3f, model4f, model5f)
realtable = anova(model1r, model2r, model3r, model4r, model5r)

tableprint = matrix(NA, nrow = 10, ncol = 8)
tableprint[1:5, 1] = "Pseudo"
tableprint[6:10, 1] = "Real"
tableprint[, 2] = c("Intercept Only", "Random Intercept", "Adjustor Variable", "Main Effects", "Interactions")
tableprint[, 3] = c(faketable$df, realtable$df)
tableprint[, 4] = apa(c(faketable$AIC, realtable$AIC), 2)
tableprint[, 5] = apa(c(faketable$BIC, realtable$BIC), 2)
tableprint[, 6] = apa(c(faketable$logLik, realtable$logLik), 2)
tableprint[, 7] = apa(c(faketable$L.Ratio, realtable$L.Ratio), 2)
tableprint[, 8] = printnum(c(faketable$p, realtable$p), gt1 = FALSE, zero = FALSE, digits = 3)

apa_table.latex(tableprint, 
          align = c("l", rep("c", 7)), 
          caption = "Area under curve model statistics",
          note = "AIC: Aikaike Information Criterion, BIC: Bayesian Information Criterion", 
          escape = FALSE,
          col.names = c("Word Type", "Model", "$df$", "AIC", "BIC", "$\\chi^2$", "$\\Delta\\chi^2$", "$p$"))

```

All predictor statistics are provided in an Excel sheet provided online at OSF. Participants were set as a random factor, and then letter frequency was used as an adjustor variable (see Table \@ref(tab:model-table2)). This variable was not significant for psuedowords, *b* = `r apa(summary(model3f)$tTable[2,1],2)` but was significant for real words, *b* = `r apa(summary(model3r)$tTable[2,1],2)`. For both real and pseudowords, the addition of the main effects of typing speed, hand switches, finger switches, and RSA were significant, *p* < .001. For psuedowords, expertise of typing speed was not significant, *b* = `r apa(summary(model4f)$tTable[3,1], 2)`, *t*(`r summary(model4f)$tTable[3,3]`) = `r apa(summary(model4f)$tTable[3,4],2)`, *p* = `r apa(summary(model4f)$tTable[3,5],3,F)`. Again, the main effect of speed was not significant for real words, *b* = `r apa(summary(model4r)$tTable[3,1], 2)`, *t*(`r summary(model4r)$tTable[3,3]`) = `r apa(summary(model4r)$tTable[3,4],2)`, *p* = `r apa(summary(model4r)$tTable[3,5],3,F)`. 

For psuedowords, hand switches was not significant, *b* = `r apa(summary(model4f)$tTable[4,1], 2)`, *t*(`r summary(model4f)$tTable[4,3]`) = `r apa(summary(model4f)$tTable[4,4],2)`, *p* = `r apa(summary(model4f)$tTable[4,5],3,F)`, but for real words, they were positively related to valence ratings, *b* = `r apa(summary(model4r)$tTable[4,1], 2)`, *t*(`r summary(model4r)$tTable[4,3]`) = `r apa(summary(model4r)$tTable[4,4],2)`, *p* < .001`. Finger switches were negatively related to both pseudowords, *b* = `r apa(summary(model4f)$tTable[5,1], 2)`, *t*(`r summary(model4f)$tTable[5,3]`) = `r apa(summary(model4f)$tTable[5,4],2)`, *p* < .001, and real words, *b* = `r apa(summary(model4r)$tTable[5,1], 2)`, *t*(`r summary(model4r)$tTable[5,3]`) = `r apa(summary(model4r)$tTable[5,4],2)`, *p* = `r apa(summary(model4r)$tTable[5,5],3,F)`. The main effect of RSA was present and positive for both psuedowords, *b* = `r apa(summary(model4f)$tTable[6,1], 2)`, *t*(`r summary(model4f)$tTable[6,3]`) = `r apa(summary(model4f)$tTable[6,4],2)`, *p* < .001 and real words, *b* = `r apa(summary(model4r)$tTable[6,1], 2)`, *t*(`r summary(model4r)$tTable[6,3]`) = `r apa(summary(model4r)$tTable[6,4],2)`, *p* < .001. These results mimics Experiment 1, except that hand switches were not significant for psuedowords in this experiment. 

## Interactions

```{r interactions2, include = FALSE}
####fake speed:lrswitch:fingerswitch:rha####
##examine speeds 
fake$lowspeed = fake$speed + sd(fake$speed)
fake$highspeed = fake$speed - sd(fake$speed)

model5fl = lme(rating ~ letterfreq + lowspeed*lrswitch*fingerswitch*rha, 
               data = fake, 
               method = "ML", 
               na.action = "na.omit",
               random = ~1|partno)
model5fh = lme(rating ~ letterfreq + highspeed*lrswitch*fingerswitch*rha, 
               data = fake, 
               method = "ML", 
               na.action = "na.omit",
               random = ~1|partno)

#summary(model5fl)
#summary(model5f)
#summary(model5fh)

##break down average and high speeds
##left right switching
fake$lowlrswitch = fake$lrswitch + sd(fake$lrswitch)
fake$highlrswitch = fake$lrswitch - sd(fake$lrswitch)

model5fal = lme(rating ~ letterfreq + speed*lowlrswitch*fingerswitch*rha, 
              data = fake, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
model5fah = lme(rating ~ letterfreq + speed*highlrswitch*fingerswitch*rha, 
              data = fake, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)

#summary(model5fal)
#summary(model5f)
#summary(model5fah)

model5fhl = lme(rating ~ letterfreq + highspeed*lowlrswitch*fingerswitch*rha, 
                data = fake, 
                method = "ML", 
                na.action = "na.omit",
                random = ~1|partno)
model5fhh = lme(rating ~ letterfreq + highspeed*highlrswitch*fingerswitch*rha, 
                data = fake, 
                method = "ML", 
                na.action = "na.omit",
                random = ~1|partno)

#summary(model5fhl)
#summary(model5fh)
#summary(model5fhh)

##average speed, low LR break down by finger switch
fake$lowfingerswitch = fake$fingerswitch + sd(fake$fingerswitch)
fake$highfingerswitch  = fake$fingerswitch - sd(fake$fingerswitch)

##avg speed low LR low FS
model5fall = lme(rating ~ letterfreq + speed*lowlrswitch*lowfingerswitch*rha, 
                data = fake, 
                method = "ML", 
                na.action = "na.omit",
                random = ~1|partno)
##avg speed low LR avg FS
model5fala = lme(rating ~ letterfreq + speed*lowlrswitch*fingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
##avg speed low LR high FS
model5falh = lme(rating ~ letterfreq + speed*lowlrswitch*highfingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
#summary(model5fall)
#summary(model5fala)
#summary(model5falh)

##avg speed avg LR low FS
model5faal = lme(rating ~ letterfreq + speed*lrswitch*lowfingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
##avg speed avg LR high FS
model5faah = lme(rating ~ letterfreq + speed*lrswitch*highfingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)

#summary(model5faal)
#summary(model5f)
#summary(model5faah)

##avg speed high LR low FS
model5fahl = lme(rating ~ letterfreq + speed*highlrswitch*lowfingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
##avg speed high LR avg FS
model5faha = lme(rating ~ letterfreq + speed*highlrswitch*fingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
##avg speed high LR high FS
model5fahh = lme(rating ~ letterfreq + speed*highlrswitch*highfingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)

#summary(model5fahl)
#summary(model5faha)
#summary(model5fahh)

##high speed avg LR low FS
model5fhal = lme(rating ~ letterfreq + highspeed*lrswitch*lowfingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
##high speed avg LR avg FS
model5fhaa = lme(rating ~ letterfreq + highspeed*lrswitch*fingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
##high speed avg LR high FS
model5fhah = lme(rating ~ letterfreq + highspeed*lrswitch*highfingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)

#summary(model5fhal)
#summary(model5fhaa)
#summary(model5fhah)

##high speed high LR low FS
model5fhhl = lme(rating ~ letterfreq + highspeed*highlrswitch*lowfingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
##high speed high LR avg FS
model5fhha = lme(rating ~ letterfreq + highspeed*highlrswitch*fingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
##high speed high LR high FS
model5fhhh = lme(rating ~ letterfreq + highspeed*highlrswitch*highfingerswitch*rha, 
                 data = fake, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)

#summary(model5fhhl)
#summary(model5fhha)
#summary(model5fhhh)

####real: lrswitch:fingerswitch:rha####
##break down by left right
real$lowlrswitch = real$lrswitch + sd(real$lrswitch)
real$highlrswitch = real$lrswitch - sd(real$lrswitch)

##low LR 
model5rl = lme(rating ~ letterfreq + speed*lowlrswitch*fingerswitch*rha, 
                 data = real, 
                 method = "ML", 
                 na.action = "na.omit",
                 random = ~1|partno)
summary(model5rl)

##avg LR
summary(model5r)

##high LR
model5rh = lme(rating ~ letterfreq + speed*highlrswitch*fingerswitch*rha, 
               data = real, 
               method = "ML", 
               na.action = "na.omit",
               random = ~1|partno)
summary(model5rh)

##break down FS
real$lowfingerswitch = real$fingerswitch + sd(real$fingerswitch)
real$highfingerswitch = real$fingerswitch - sd(real$fingerswitch)

##avg LR low FS
model5al = lme(rating ~ letterfreq + speed*lrswitch*lowfingerswitch*rha, 
               data = real, 
               method = "ML", 
               na.action = "na.omit",
               random = ~1|partno)
summary(model5al)

##avg LR avg FS
summary(model5r)

##avg LR high FS
model5ah = lme(rating ~ letterfreq + speed*lrswitch*highfingerswitch*rha, 
               data = real, 
               method = "ML", 
               na.action = "na.omit",
               random = ~1|partno)
summary(model5ah)

##high LR low FS
model5hl = lme(rating ~ letterfreq + speed*highlrswitch*lowfingerswitch*rha, 
               data = real, 
               method = "ML", 
               na.action = "na.omit",
               random = ~1|partno)
summary(model5hl)

##high LR avg FS
model5ha = lme(rating ~ letterfreq + speed*highlrswitch*fingerswitch*rha, 
               data = real, 
               method = "ML", 
               na.action = "na.omit",
               random = ~1|partno)
summary(model5ha)

##high LR high FS
model5hh = lme(rating ~ letterfreq + speed*highlrswitch*highfingerswitch*rha, 
               data = real, 
               method = "ML", 
               na.action = "na.omit",
               random = ~1|partno)
summary(model5hh)

```

Here's where I tested if experiment 2 was just a fluke of the 3-4 letter words used, and alas, it probably is. 
```{r testing, include = FALSE}

tempexp2 = subset(exp1data, trialcode %in% exp2data$trialcode)
tempexp2$rha = scale(tempexp2$rha, scale = F)
tempexp2$fingerswitch = scale(tempexp2$fingerswitch, scale = F)
tempexp2$lrswitch = scale(tempexp2$lrswitch, scale = F)
tempexp2$speed = scale(tempexp2$speed, scale = F)

real = subset(tempexp2, real_fake == "real")
fake = subset(tempexp2, real_fake == "fake")

####intercept only model####
model1f = gls(rating ~ 1, 
             data = fake, 
             method = "ML", 
             na.action = "na.omit")
summary(model1f)

model1r = gls(rating ~ 1, 
             data = real, 
             method = "ML", 
             na.action = "na.omit")
summary(model1r)

####random intercept only model####
model2f = lme(rating ~ 1, 
             data = fake, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model2f)

model2r = lme(rating ~ 1, 
             data = real, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model2r)


####controls predictor models####
model3f = lme(rating ~ letterfreq, 
              data = fake, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
summary(model3f)

model3r = lme(rating ~ letterfreq, 
             data = real, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model3r)

####main effects models####
model4f = lme(rating ~ letterfreq + speed + lrswitch + fingerswitch + rha, 
             data = fake, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model4f)

model4r = lme(rating ~ letterfreq + speed + lrswitch + fingerswitch + rha, 
             data = real, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model4r)

####interaction models####
model5f = lme(rating ~ letterfreq + speed*lrswitch*fingerswitch*rha, 
              data = fake, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
summary(model5f)

model5r = lme(rating ~ letterfreq + speed*lrswitch*fingerswitch*rha, 
              data = real,
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
summary(model5r)
```





lrswitch:fingerswitch:rha FOR REAL WORDS 