---
title             : "An Extension of the QWERTY Effect: Not Just the Right Hand, Expertise and Typability Predict Valence Ratings of Words"
shorttitle        : "QWERTY EFFECT EXTENSION"

author: 
  - name          : "Erin M. Buchanan"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "901 S. National Ave, Springfield, MO 65897"
    email         : "erinbuchanan@missouristate.edu"
  - name          : "Kathrene D. Valentine"
    affiliation   : "2"

affiliation:
  - id            : "1"
    institution   : "Missouri State University"
  - id            : "2"
    institution   : "University of Missouri"

author_note: |
  Erin M. Buchanan is an Associate Professor of Quantitative Psychology at Missouri State University. K. D. Valentine is a Ph.D. candidate at the University of Missouri. 
  
abstract: |
  Typing is a ubiquitous daily action for many individuals; yet, research on how these actions have changed our perception of language is limited. The QWERTY effect is an increase in valence ratings for words typed more with the right hand on a traditional keyboard (Jasmin & Casasanto, 2012). Although this finding is intuitively appealing given both right handed dominance and the smaller number of letters typed with the right hand, extension and replication of the right side advantage is warranted. The present paper reexamined the QWERTY effect within the embodied cognition framework (Barsalou, 1999) and found that the right side advantage is replicable to new valence stimuli, as well as experimental manipulation. Further, when examining expertise, right side advantage interacted with typing speed and typability (i.e. alternating hand keypresses or finger switches) portraying that both skill and our procedural actions play a role in judgment of valence on words.
  
keywords          : "keyboard, valence, QWERTY, word norms"

bibliography      : ["q_bib.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages, include = FALSE}
library("papaja")
library(nlme)
library(MOTE)
p.value = function(x){
  if (x < .001) { return("< .001")}
  else { return(apa(x, 3, F))}
}
```

From its creation in 1868, to its appearance in our homes today, the QWERTY keyboard has held the interest of psychologists. The process of typing on a keyboard requires many procedures to function in tandem, which creates a wealth of actions to research [@Inhoff1997]. @Rumelhart1982's computer model of skilled typing is still highly influential. They hypothesize that typing results from the activation of three levels of cognition: the word level, the keypress level, and the response level. They believe that after word perception, the word level is activated, causing the keypress level to initiate a schema of the letters involved in typing the word. This schema includes the optimal position on the keyboard for that specific hand-finger combination to move to at the appropriate time for individual keystrokes. Concurrently, the response system sends feedback information to initiate a keypress motion when the finger is in the appropriate space. Their theory proposes that schemata and motion activations occur simultaneously, constantly pulling or pushing the hands and fingers in the right direction.

While many studies have focused on errors in typing to investigate response system feedback [@Logan1999], @Logan2003 argued for parallel activation of keypresses. He examined the Simon effect to show more than one letter is activated at the same time, and consequently, the second keypress motion is begun before the first keypress is done. The Simon effect occurs when congruent stimuli create faster responses than incongruent stimuli, much like the Stroop task [@Simon1990; @Simon1969]. For example, if we are asked to type the letter f (a left handed letter), we type it faster if the f is presented on the left side of the screen. Similarly, @Rieger2004 reported finger-congruency effects by altering a Stroop task: participants were required to respond to centrally presented letters based on color-key combinations. When the letter and color were congruent (i.e. a right-handed letter was presented in the designated color for a right response), the skilled typists' responses were faster than incongruent combinations. Further, this effect was present when participants responded to items with their hands crossed on the responding device, suggesting the effect was expertise-based rather than experiment-response based. These results imply that automatic actions stimulate motor and imagery representations concurrently and may be linked together in the brain [@Hommel2001; @Logan1998; @Rieger2004]. This dual activation of motor and imagined items is the basis for embodied cognition, a rapidly expanding field in psychology [@Barsalou1999; @Salthouse1986]. 

## Embodied Cognition

While the mind was traditionally considered an abstract symbol processor [@Newell1976], newer cognitive psychology theories focus on the interaction between the brain's sensorimotor systems and mental representations of events and objects [@Barsalou1999; @Zwaan1999]. The interplay between these systems has been found in both neurological [@Hauk2004; @Lyons2010; @Tettamanti2005] and behavioral research [@Cartmill2012; Holt2006; @Zwaan2006]. Motor representations of tasks are activated even when not specifically asked to perform the task, and if the action is well-learned, the task is perceived as pleasant [@Beilock2007; @Ping2009; @Yang2009]. For example, @Beilock2007 asked novice and expert typists to pick which one of two letter dyads they preferred, which were either different hand combinations (CJ) or same finger combinations (FV). They found that novices have no preference in selection, while expert typists more reliably picked the combinations that were easier to type. To show that this effect was due to covert motor representation activation, and thus, expanding on findings from @VandenBergh1990, participants also made preference selections while repeating a keypress combination. When expert motor planning was distracted by remembering the pattern presented, no preference for letter dyads was found, indicating that the simultaneous activation of the motor representation was necessary to influence their likability ratings. Similar embodied findings have also been portrayed with emotionally charged sentences and facial movements [@Havas2007], positive-negative actions, such as head nodding or arm movements [@Glenberg2009; @Ping2009], and perceptuomotor fluency [@Oppenheimer2008; @Yang2009].

## Body Specificity Hypothesis
Using an embodied framework, @Casasanto2009 has proposed that handedness dictates preference because our representations of actions are grounded in our physical interactions with the environment. In several studies, he portrayed that handedness influenced preference for spatial presentation (i.e. left handed individuals associate “good” with left, while right handed individuals associate “good” with right), which in turn influenced judgments of happiness and intelligence and our decision making in hiring job candidates and shopping. In all these studies, participants reliably selected the hand-dominant side more often, which does not match cultural or neurolinguistic representations of positive-is-right and negative-is-left [@Davidson1992]. These findings imply that our handedness is a motor expertise that causes ease of action on the dominant side to positively influence our perceptions of items presented on that side. Further, @Casasanto2011 compiled a review of body specific actions and their representation in the brain using fMRIs. Handedness interacted with imagining actions, reading action, and perceiving the meanings of action verbs, such that fMRI patterns were mirrored for left and right handed participants matching their dominant side.

## The QWERTY Effect

These effects inspired @Jasmin2012 to propose the idea that typing, an action that often replaces speaking, has the ability to create semantic changes in how we perceive words. The asymmetrical arrangement of letters on the QWERTY keyboard increases fluency of typing letters on the right side because there are fewer keys, and thus, less competition for fingers. That arrangement should then cause us to perceive the letters on the right side as more positive and letters on the left side as more negative. Consequently, words that are composed of more letters from the right side (the right side advantage; RSA) should be rated as more positive than those with more letters on the left. They found this preference for RSA over three languages (English, Spanish, and Dutch), and the effect was even stronger on words created after the invention of the QWERTY keyboard (i.e. lol), as well as evident in pseudowords such as plook. However, in contrast to the body specificity hypothesis, left and right handed participants showed the same trend in effects for positive-is-right words. 

# Current Study

The current study examined the right side advantage's interaction with traditional embodied cognition definitions (expertise, fluency). We analyzed the different implications of the body specificity hypothesis and a more general embodied hypothesis by testing the following: 1) To examine embodied cognition, we coded each word for number of hand alternations (akin to @Beilock2007's different hand preferences). Given that typing involves the procedural action system, we would also expect to find that increased hand switches are positively related to ratings of valence because words that are typed on alternating hands are easier to type. 2) The interaction between RSA and switches was examined to determine if these hypotheses can be combined (i.e. we only like right handed words because we have to switch back and forth to type the more commonly used letters, such as *e* or *a*).

# Experiment 1

# Method

## Participants

```{r datascreen, include=FALSE}
####import files####
exp1data = read.csv("exp1_final.csv")

####data screening####
##accuracy checks
summary(exp1data)
part = length(table(exp1data$partno))
#table(exp1data$speed)
#hist(exp1data$speed)
exp1data$real_fake = factor(exp1data$real_fake,
                            levels = c("fake", "real"))

##missing data
##no missing data to replace
##figure out the couple missing data points
rows = nrow(exp1data)
parttable = as.data.frame(table(exp1data$partno))
#parttable[parttable$Freq > 120, ]
missdatapoints = nrow(parttable[parttable$Freq < 120, ])
missdatapoints119 = nrow(parttable[parttable$Freq == 119, ])

##outliers
#with(exp1data, plot(speed, rating))
mahal = mahalanobis(exp1data[ , c(5,7)],
                    colMeans(exp1data[ , c(5,7)], na.rm = T),
                    cov(exp1data[ , c(5,7)], use = "pairwise.complete.obs"))
cutoff = qchisq(1-.001, ncol(exp1data[ , c(5,7)]))
table(mahal < cutoff)
exp1data$partno[ mahal > cutoff] ##all same participant

##assumptions
correlations = cor(exp1data[, c(5,7, 13, 14, 15, 16, 17)], use="pairwise.complete.obs")
correlations
symnum(correlations)
#do not use length and fingerswitch together 

##make the random stuff
random = rchisq(nrow(exp1data), 7)
#fake = lm(random~., data=exp1data[ , c(5, 7, 12, 13, 14, 15, 17)])
fake = lm(random~., data=exp1data[ , c(5, 7)])

##get the linearity plot
##create the standardized residuals
standardized = rstudent(fake)
{qqnorm(standardized)
abline(0,1)}

##multivariate normality
hist(standardized, breaks=15)

##homogeneity and homoscedaticity
fitvalues = scale(fake$fitted.values)
{plot(fitvalues, standardized) 
abline(0,0)
abline(v = 0)}

exp1data$hand[exp1data$hand == 0] = NA
exp1data$hand[exp1data$hand == "the"] = NA
exp1data$hand[exp1data$hand == "left"] = "Left"
exp1data$hand[exp1data$hand == "right"] = "Right"
exp1data$hand[exp1data$hand == "RIGHT"] = "Right"
exp1data$hand = droplevels(exp1data$hand)

temp = exp1data[!duplicated(exp1data$partno), ]
hand = table(temp$hand) / length(temp$hand) * 100

```

Participants (*N* = `r part`) were recruited from the university undergraduate human subject pool and received course credit for their time. `r rows` rows of data were present for these participants, where `r (part - missdatapoints)` participants included complete data (i.e. 120 rows, see below), `r missdatapoints119` were missing one data point, and `r (missdatapoints - missdatapoints119)` were missing many data points. All data points were included, and missing data points were usually computer error (i.e. freezing during the experiment) or participant error (i.e. missed key press). 

Rating data were screened for multivariate outliers, and one participant's ratings were found to have extreme Mahalanobis distance scores [@Tabachnick2012] but were kept in the data set. `r apa(hand['Left'], 1)` percent of the sample was left-handed, `r apa(hand['right_left'], 1)` marked ambigdextrious, and `r apa(100 - sum(hand), 1)` was missing handedness information. The average typing speed was `r apa(mean(temp$speed), 2)` (*SD* = `r apa(sd(temp$speed), 2)`, and the average percent accuracy rate for the typing test was `r apa(mean(temp$error), 2)` (*SD* = `r apa(sd(temp$error), 2)`).

## Materials

Both Experiment 1 and Experiment 2 use the English ANEW [@Bradley1999] norms to create stimuli for this study, in an effort to replicate @Jasmin2012 experiments, and 240 words were selected for this experiment (120 real words, 120 pseudowords). Pseudowords were selected from Appendix E of the supplementary materials presented from the QWERTY publication. These words were coded as described below for RSA, switches, word length, and letter frequency. Average word length was `r apa(mean(exp1data$length), 2)` (*SD* = `r apa(sd(exp1data$length), 2)`; range = `r min(exp1data$length)` - `r max(exp1data$length)`).

## Coding

Each of the words used in this experiment and Experiment 2 were coded for control and experimental variables. Control variables included word length and average letter frequency. Average letter frequency was created by averaging the English letter frequency  [@Lewand2000] for each letter in a word. Words with high average letter frequencies contain more commonly used letters (*e, t, a, o*); while words with lower frequencies use more of the less common letters (*z, q, x, j*). Experimental variables included RSA, number of hand switches, and number of finger switches. Typing manuals were consulted, and letters were coded as left (*q, w, e, r, t, a, s, d, f, g, z, x, c, v, b*) or right-handed letters (*y, u, i, o, p, h, j, k, l, n, m*). Left handed letters were coded with -1 and right handed letters with +1, which created summed scores indicating the overall right side advantage for a word. Words were coded for the number of hand switches within a word using the left-right coding system described above. Finally, the number of finger switches were coded using traditional typing manuals for each finger. Finger switches was highly correlated with word length, *r* = `r apa(cor(exp1data$length, exp1data$fingerswitch), 2, F)`, and therefore, word length was excluded as a control variable due the interest in typing skill for experimental hypotheses.

## Procedure

Upon consent to participate in the experiment, participants were given a typing test by using a free typing test website [@Inc2013]. Each participant typed Aesop's Fables for one minute before the website would reveal their typing speed and accuracy rate, which was recorded by the experimenter. After this test, participants indicated their dominant writing hand. Participants were then given 120 of the 240 stimuli to rate for pleasantness (60 real words, 60 pseudowords). This smaller number of stimuli was used to control fatigue/boredom on participants. These stimuli were counterbalanced across participants, and the order of the stimuli was randomized. Participants were told to rate each word for how pleasant it seemed using a 9 point Likert type scale (1 - *very unpleasant*, 4 - *neutral*, 9 - *very pleasant*). The same self-assessment manikin from @Jasmin2012 was shown to participants at the top of the computer screen to indicate the points on the Likert scale. The words appeared in the middle of the screen in 18 point Arial font. Participants then typed the number of their rating on the computer keyboard. Once they rated all stimuli, participants were debriefed and allowed to leave.

# Results

## Data Analytic Plan

Because each participant constituted multiple data points within the dataset, a multilevel model was used to control for correlated error [@Gelman2006]. @Pinheiro2017's *nlme* package in *R* was used to calculate these analyses. A maximum likelihood multilevel model was used to examine hypotheses of interactions between typing speed, hand/finger switching, and RSA while controlling for letter frequency predicting item pleasantness ratings. Participants were included as a random intercept factor. Typing speed, finger/hand switches, and RSA were mean centered before analyses to control for multicollinearity.

```{r setupexp1, include=FALSE}
####separate real fake####
real = subset(exp1data,real_fake == "real")
fake = subset(exp1data,real_fake == "fake")

####centering the variables####
real$speed = scale(real$speed, scale = F)
real$lrswitch = scale(real$lrswitch, scale = F)
real$fingerswitch = scale(real$fingerswitch, scale = F)
real$rha = scale(real$rha, scale = F)

fake$speed = scale(fake$speed, scale = F)
fake$lrswitch = scale(fake$lrswitch, scale = F)
fake$fingerswitch = scale(fake$fingerswitch, scale = F)
fake$rha = scale(fake$rha, scale = F)
```

## Main Effects

```{r meexp1, include=FALSE}
 ####intercept only model####
model1f = gls(rating ~ 1, 
             data = fake, 
             method = "ML", 
             na.action = "na.omit")
summary(model1f)

model1r = gls(rating ~ 1, 
             data = real, 
             method = "ML", 
             na.action = "na.omit")
summary(model1r)

####random intercept only model####
model2f = lme(rating ~ 1, 
             data = fake, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model2f)

model2r = lme(rating ~ 1, 
             data = real, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model2r)


####controls predictor models####
model3f = lme(rating ~ letterfreq, 
              data = fake, 
              method = "ML", 
              na.action = "na.omit",
              random = ~1|partno)
summary(model3f)

model3r = lme(rating ~ letterfreq, 
             data = real, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model3r)

####main effects models####
model4f = lme(rating ~ letterfreq + speed + lrswitch + fingerswitch + rha, 
             data = fake, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model4f)

model4r = lme(rating ~ letterfreq + speed + lrswitch + fingerswitch + rha, 
             data = real, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model4r)
```


## Interactions

```{r interactexp1, include = FALSE}
####interaction models####
model5f = lme(rating ~ letterfreq + speed*lrswitch*fingerswitch*rha, 
             data = fake, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model5f)

model5r = lme(rating ~ letterfreq + speed*lrswitch*fingerswitch*rha, 
             data = real, 
             method = "ML", 
             na.action = "na.omit",
             random = ~1|partno)
summary(model5r)

faketable = anova(model1f, model2f, model3f, model4f, model5f)
realtable = anova(model1r, model2r, model3r, model4r, model5r)

tableprint = matrix(NA, nrow = 10, ncol = 8)
colnames(tableprint) = c("Word Type", "Model", "df", "AIC", "BIC", "logLik", "X2", "p")
tableprint[1:5, 1] = "Fake"
tableprint[6:10, 1] = "Real"
tableprint[, 2] = c("Intercept Only", "Random Intercept", "Control Variables", "Main Effects", "Interactions")
tableprint[, 3] = c(faketable$df, realtable$df)
tableprint[, 4] = apa(c(faketable$AIC, realtable$AIC), 2)
tableprint[, 5] = apa(c(faketable$BIC, realtable$BIC), 2)
tableprint[, 6] = apa(c(faketable$logLik, realtable$logLik), 2)
tableprint[, 7] = apa(c(faketable$L.Ratio, realtable$L.Ratio), 2)
tableprint[, 8] = c(faketable$p, realtable$p)
#tableprint[, 8] = apply(tableprint[, 8], 1, p.value)
```


# Experiment 2

# Method

## Participants

```{r datascreen2}
exp2data = read.csv("exp2_final.csv")
```


Similar to Experiment 1, sixty participants were recruited from the university undergraduate human subject pool and received course credit for their time. Rating data were screened for multivariate outliers. Again, one participant's ratings were found to have extreme Mahalanobis distance scores [@Tabachnick2012]. However, this individual's ratings were left in the data set. Approximately 8 percent of the sample was left-handed. The average percent accuracy rate for the typing test was 93.58 (*SD* = 5.26). 

## Procedure

While materials and coding were the same for Experiments 1 and 2, procedure for Experiment 2 differed slightly. In this study, when participants were shown the word (or pseudoword) on the screen, they were first asked to type the word on the keyboard in front of them. After they had typed the word, they were then asked to rate the word for pleasantness using the scale and self-assessment manikin discussed previously.  

# Results

## Main Effects

## Interactions

# Discussion

YADA SCHMADA CHANGE THIS SECTION
These results imply that the QWERTY keyboard has influenced our perceptions of words, in a more complex way than a simple body specificity hypothesis. In the overall normed database analyses, the original QWERTY effect was replicable across a large body of various types of stimuli (verbs, Twitter, category norms), with much the same size of effect as @Jasmin2012 published. Word length was often negatively related to valence ratings, which indicated that we like shorter words to type. Average letter frequency was usually a positive predictor of valence ratings wherein ratings are higher for words with more frequent letters; however, these effects were inconsistent. Our measure of fluency (switches) varied across stimulus sets but it appears, by analyzing multiple sources of ratings for words at the same time, that there might have been an interaction between RSA and number of switches. This interaction portrayed that we find words that switch off of left-handed keypresses as more pleasant, while right-handed keypresses are preferable by switching hands less often.
These effects were examined in more detail in Experiment 2, which incorporated @Beilock2007 study by including typing speed as a measure of expertise. Word ratings turned out to be quite complex with a four-way interaction between real/pseudowords, switches, RSA, and typing speed. All analyses showed a positive effect of right-side words, as well as if they were shorter and used more frequent letters. However, for pseudowords, no other effects were significant. Both @Beilock2007 and @VandenBergh1990 showed expert preferences for two and three letter combinations that were typed with different fingers. Our results could imply that our embodied actions influence preferences for procedures that are more likely in our environment. While our pseudowords were legal English phoneme combinations, they are extremely unlikely to have been previously practiced or encountered in our daily tasks. Therefore, switching preference will not extend to pseudowords (unpracticed actions) because they are not fluent [@Oppenheimer2008]. 

The effect of expertise was shown on real words, where the three-way interaction between RSA, switches, and typing speed was examined by separating out right, equal, and left-handed words. For right-handed words, typing speed (or the interaction) was not a significant predictor of valence, and while not significant, number of switches was negatively related to valence ratings. For equally right-left and left-handed words, pleasantness ratings increase by switching back and forth to the right hand. Further, left-handed words showed an interaction between our two embodied cognition variables, where the number of switches increases valence ratings as the typing speed of the participant decreases. Therefore, it appears that as participants gain fluency through increased typing speed, the number of switches back and forth for left-handed words matters less for pleasantness ratings. Many of the most frequent letters on the QWERTY keyboard are on the left side, which may frustrate a slow typist because of the need to coordinate finger press schemata that involve same finger muscle movements [@Rumelhart1982]. Consequently, the number of switches becomes increasingly important to help decrease interference from the need to continue to use the same hand. The ease of action by switching back and forth is then translated as positive feelings for those fluent actions [@Oppenheimer2008].

These embodied results mirror a clever set of studies by @Holt2006 wherein they showed participants sentences that matched or did not match a set of pictures (i.e. the umbrella is in the air paired with a picture of an open umbrella). Given dual-coding theory [@Paivio1991], it was not surprising that participants were faster to indicate picture-sentence matches than non-matches [also see @Stanfield2001; @Zwaan2002]. Further, they showed these results extended to an expertise match; hockey and football players were much faster for sentence-picture combinations that matched within their sport than non-matches, while novices showed no difference in speed for matches or non-matches on sports questions. Even more compelling are results that these effects extend to fans of a sport and are consistent neurologically [i.e. motor cortex activation in experts; @Beilock2008]. These studies clearly reinforce the idea that expertise and fluency unconsciously affect our choices, even when it comes to perceived pleasantness of words. 

This extension of the QWERTY effect illuminates the need to examine how skill can influence cognitive processes. Additionally, typing style, while not recorded in this experiment, could potentially illuminate differences in ratings across left-handed and right-handed words. Hunt-and-peck typists are often slower than the strict typing manual typists, which may eliminate or change the effects of RSA and switches since typists may not follow left or right hand rules and just switch hands back and forth regardless of key position. The middle of a QWERTY layout also poses interesting problems, as many typists admit to “cheating” the middle letters, such as t, and y or not even knowing which finger should actually type the b key. Further work could also investigate these effects on other keyboard layouts, such as Dvorak, which was designed to predominately type by alternating hands to increase speed and efficiency [@Noyes1983]. 

\newpage

# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
